---
title: "Hunayn"
author: "Kieran Patel"
date: "10/12/2017"
output: html_document
---


####Hunayn####


```{r}
M <- matrix(0, nrow = 7, ncol = 7)
M [1, 2] <- 12
M [1, 3] <- 10
M [1, 4] <- 18
M [2, 3] <- 5
M [2, 5] <- 9
M [3, 4] <- 9
M [3, 5] <- 10
M [3, 6] <- 12
M [4, 6] <- 12
M [5, 6] <- 10
M [5, 7] <- 7
M [6, 7] <- 25
M

f.matrix <- matrix(0, nrow = 7, ncol = 7)

f.matrix [1, 2] <- 8
f.matrix [1, 3] <- 5
f.matrix [1, 4] <- 7
f.matrix [2, 3] <- 4
f.matrix [2, 5] <- 4
f.matrix [3, 4] <- 0
f.matrix [3, 5] <- 9
f.matrix [3, 6] <- 0
f.matrix [4, 6] <- 7
f.matrix [5, 6] <- 10
f.matrix [5, 7] <- 3
f.matrix [6, 7] <- 17
f.matrix

f.matrix. <- f.matrix
f.matrix. [1, 1] <- 20
f.matrix. [7, 7] <- 20
f.matrix.

#adj.matrix <- matrix(0, nrow = 7, ncol = 7)
#adj.matrix [1, 2] <- 1
#adj.matrix [1, 3] <- 1
#adj.matrix [1, 4] <- 1
#adj.matrix [2, 3] <- 1
#adj.matrix [2, 5] <- 1
#adj.matrix [3, 4] <- 1
#adj.matrix [3, 5] <- 1
#adj.matrix [3, 6] <- 1
#adj.matrix [4, 6] <- 1
#adj.matrix [5, 6] <- 1
#adj.matrix [5, 7] <- 1
#adj.matrix [6, 7] <- 1

#adj.matrix [2, 1] <- 1
#adj.matrix [3, 1] <- 1
#adj.matrix [4, 1] <- 1
#adj.matrix [3, 2] <- 1
#adj.matrix [5, 2] <- 1
#adj.matrix [4, 3] <- 1
#adj.matrix [5, 3] <- 1
#adj.matrix [6, 3] <- 1
#adj.matrix [6, 4] <- 1
#adj.matrix [6, 5] <- 1
#adj.matrix [7, 5] <- 1
#adj.matrix [7, 6] <- 1
#adj.matrix


adj.matrix <- M
for (i in 1:7){
  for (j in 1:7){
    if (adj.matrix[i, j] > 0){
      adj.matrix [i, j] <- 1
      }
  }
}
adj.matrix
```

##A
```{r}
list.of.edges <- which(adj.matrix == 1, arr.ind = TRUE)
list.of.edges
list.of.edges. <- list.of.edges[sort.list(list.of.edges[, 1]), ]
list.of.edges.
```
This would be useful with the Ford-Fulkerson algorithm as we need to know which paths are possible to add to the route.


##B
```{r}
#f.matrix. <- f.matrix
#f.matrix. [1, 1] <- 20
#f.matrix. [7, 7] <- 20

rows <- f.matrix.[1:7, 2:7]
rows
columns <- f.matrix.[1:6, 1:7]
columns



feasibility <- function(M, f.matrix, f.matrix., adj.matrix){
  i <- 1
  for(i in 1:7){
  if(all(f.matrix <= M) & (all(sum(rows[i, ]) == sum(columns[, i]))) & (f.matrix.[1, 1] == f.matrix.[7, 7])){
    return(print('feasible'))
    }
  else {
    return(print('not feasible'))
  }
    i <- i + 1
    }
}

feasibility(M, f.matrix, adj.matrix)
```

I have two conditions to check for feasibility so I will put them inside two separate if statements and then check that for any if statements which return 'feasible', there must be two 'feasible' returned for the flow to be be feasible.

I then googled how to add multiple conditions to an if statement and the & did this. I then remembered that the flow in must equal the flow out so there was another final condition.

I was not sure how to fully implement the conditions as the 7 x 7 matrix has some comlexities as we need to check rows 1:7 for the matrix f.matrix[1:7, 2:7] equal columns 1:6 for matrix f.matrix[1:6, 1:7].

I then changed the values of the code to reflect these values.

I then realised that the flow was still not feasible as I added this to my code
f.matrix [1, 1] <- 20
f.matrix [7, 7] <- 20
which meant that I was using a different version of the M matrix.

We need the flow to be feasible in order to use the Ford-Fulkerson algorithm. This is the first step for the algorithm.

##Ci
```{r}
n <- 7
V <- sample (1 : 7)
V
S <- c (4, 5, 6)
S_c <- V[! V %in% S]
S_c
```

This function which I found online after reasrching how to remove elements from a vector, removes elements, S, from vector V, to create vector S_c.

We need to identfy which paths are already in S and which are in S_c so we know which we can add to S depending on whether flows are < the capacities for a flow and capacity which the path is from S and goes towards S_c. This cut relies on knowing whether they are in S or S_c.

##Cii
```{r}
i <- 1
list <- list()
for(a in S_c){
  which.edge <- which(a == list.of.edges.[1,])
  list[[i]] <- which.edge
  i <- i + 1
}
list

#(1, 2), (1, 3), (2, 3)
```

I first attempted this
which.edge <- which(S_c == list.of.edges[1,], arr.ind = TRUE)
which.edge
which gave one output which I knew to be wrong as they are meant to be #(1, 2), (1, 3), (2, 3)

I then tried the following two attempted but they outputted strange answers but closer than before as I was receiving an error before.
for(i in S_c){
  which.edge <- which(i == list.of.edges.[1,])
  return(which.edge)
  
  i <- 1
list <- list()
for(i in S_c){
  which.edge <- which(i == list.of.edges.[1,])
  list[[i]] <- which.edge
  i <- i + 1
}
list

This gives us the potential paths we can take to the route we are making. This is necessary as we need to be using valid paths.

##D
```{r}
increase.flow <- min(M - f.matrix)
increase.flow
```

The flow demonstrated here cannot be increased as the increase.flow output is 0.

This ensures the flow is maximal and that it is optimized so that there is no more flow which can pass through the network.

Elements which are not discussed are whether the flow moves the other way. Another element which is not discussed is the cuts between S and S_c.