---
title: "Ptolemy"
author: "Kieran Patel"
date: "12/11/2017"
output: html_document
---


####Ptolemy Coursework####


##(a)
There are 25 1x1 squares, 16 2x2 squares, 9 3x3 squares, 4 4x4 squares, which I clearly saw by drawing out the matrices and identifying the pattern.

##(b)
```{r}
#M
Matrix.list <- list()
n <- 1
#I want to make an empt list so that the for loop fills up the list
i <- 1
j <- 1

for (i in 1:5){
  for (j in 1:5){
    M <- matrix(0, nrow = 5, ncol = 5)
    M[i:i, j:j] <- 1
    Matrix.list [[n]] <- M
    n <- n + 1}}
#Matrix.list
```

I didn't want to type out each matrix out as there are 54 and that would be laborious.

I am trying to code each of the 1x1 matrices. After googling, I found instead of for (i = [1:5], j [1:5]), I should type for (i in 1:5, j in 1:5) as I was trying to make it so for any i between 1 and 5 and any j between 1 and 5 some codes will be executed.
I am getting errors when creating my for loop to create all the 1x1 squares with the following code
for (i in 1:5, j in 1:5). The comma in this code them comes up with an error and I'm not sure how to find multiple conditions but I then realised that this is similar to a double sum so I decided to create another for loop for the js.
{A[i, j] <− 1}.

The code now doesn't have any errors but the output is not what I was hoping, it has made all the entries 1.

I then tried to print the matrix and realised it is adding the new entries of 1 onto the same matrix, so I tried to add this new matrix into a list.

Matrix.list <- list()
i <- 1
j <- 1
for (i in 1:5){
  for (j in 1:5){
    M <- matrix(0, nrow = 5, ncol = 5)
    M[i, j] <- 1
    Matrix.list <- M}}
    #print(Matrix.list)
    #print(M)}}
Matrix.list

This just outputs the last matrix of the loop of 1x1 squares.

The curved brackets did not work but a helper during the computing class said to use double square brackets but this still didn't work since the entry in the double brackets is blank and it needs to be defined and we'll call this n for which number the matrix is. Also for completeness, I coded the square matrix of 1 to be from i:i and j:j.

Matrix.list now outputs all the 1x1 squares but cannot pick out certain elements of the list, I assume this is because the inputs of the list are not labelled as 1, 2, 3, etc. I did this by introducing each new matrix as the (n+1)th entry in the list.

```{r}
n <- 26
#I now changed n to 26 so that the entries in the list are not overwritten.
i <- 1
j <- 1
for (i in 1:4){
  for (j in 1:4){
    M <- matrix(0, nrow = 5, ncol = 5)
    M[i: (i + 1), j: (j + 1)] <- 1
    Matrix.list [[n]] <- M
    n <- (n + 1)}}
```

```{r}
n <- 42
i <- 1
j <- 1
for (i in 1:3){
  for (j in 1:3){
    M <- matrix(0, nrow = 5, ncol = 5)
    M[i: (i + 2), j: (j + 2)] <- 1
    Matrix.list [[n]] <- M
    n <- (n + 1)}}
```

```{r}
n <- 51
i <- 1
j <- 1
for (i in 1:2){
  for (j in 1:2){
    M <- matrix(0, nrow = 5, ncol = 5)
    M[i: (i + 3), j: (j + 3)] <- 1
    Matrix.list [[n]] <- M
    n <- (n + 1)}}
print(Matrix.list)
```


I then edited some of the code to adjust for the now 2x2 squares and then added this onto the list but now beginning as the 26th entry. I then realised something similar to my last coursework where I had to bracket the i+1 because it confuses R and doesn't compute the code in the way that I want it.

The 2x2 squares are outputting but my code for the 1 to 25th entry of the list have now become 'NULL'. I tried to clear my environment and then run all my code again and it worked.

##(c)
```{r}
vector.list <- list()
n <- 1
for (n in 1:54){
  vector.list [[n]] <- as.vector(t(Matrix.list[[n]]))
  n <- n + 1}
#vector.list

A <- matrix(unlist(vector.list), ncol = 25, byrow = TRUE)
A. <- t(A)
A.
```

I remembered from Statistics last year we can easily turn a matrix into a vector with the as.vector function and I put this in a for loop.

I am trying to create a matrix with all the constraints which are currently in the columns of the vector.list.

M. <- matrix(0, ncol = 1, nrow = 54)
for (n in 1:54){
  M. [n] <- as.vector(t(Matrix.list[[n]]))
  n <- n + 1}
  
This creates just the first constraint when I was hoping to have the matrix of constraints.
I tried several attempts to create this matrix, here are my attempts

for (n in 1:54){
  M.[ [1:25], n] <- vector.list[[n]]
 n <- n + 1}

for (m in 1:54){
  M. <- cbind(vector.list[1:54])
  m <- m + 1}
  
  for (i in 1:54){
  for (j in 1:25){
  M. [n] <- as.vector(t(Matrix.list[[n]]))
  n <- n + 1}
  
I then researched into turning a list into a matrix and found an easy function 'unlist'.

```{r}
b <- rep(1, 54)
#54 for 54 objective variables
c <- rep(1, 25)
ineqs <- rep("=", 25)
#25 for 25 constraints
```

##(d)
```{r}
library(lpSolve)
lp("min", b, A., ineqs, c)$solution
```

The solution for this problem, is the 3rd, 4th and 5th from the first row, 3rd from the second row,  1x1 squares
The 1st from the first row, 4th from the second row, 4th from the fourth row, 2x2 squares.
The 1st from the first row, 3x3 square.

I did not need to include any additional constraints.

##(e)
I did not need to use the int.vec parameter, nor did I get any strange non-integer solutions which is surprising as the question implies that we should.