---
title: "Hypatia_and_Theon"
author: "Kieran Patel"
date: "29/11/2017"
output: html_document
---


####Hypatia Coursework####


##(a)
Gomory cuts reduce a feasible region to include all the integer solutions as before but fewest number of non-integer solutions.

##(b)
```{r}
library (lpSolve)
f.obj1 <- c(12, 20, 15, 0, 0, 0, 0)
f.con1 <- matrix(c(2, 5, 3, 1, 0, 0, 0, 1, 1, 1, 0, 1, 0, 0, 0, 1, 0, 0, 0, -1, 1), nrow = 3, byrow = TRUE)
f.ineqs1 <- c("=", "=", '=')
f.rhs1 <- c(50, 12, 5)
lp("max", f.obj1, f.con1, f.ineqs1, f.rhs1)
lp("max", f.obj1, f.con1, f.ineqs1, f.rhs1)$solution
```

I cannot seem to get my s3 to be 2. In this question, I have to put <= and >= instead of = to get the same answer as in the notes.

I kept fiddling with my input values and realised it was because there are 4 slack variables.

I have managed to make it work, I found out that my inequality had not been changed to equality.

##(c)
```{r}
f.obj2 <- c(12, 20, 15, 0, 0, 0, 0)
f.con2 <- matrix(c(2, 5, 3, 1, 0, 0, 0, 1, 1, 1, 0, 1, 0, 0, 0, 1, 0, 0, 0, -1, 1), nrow = 3, byrow = TRUE)
f.ineqs2 <- c("=", "=", '=')
f.rhs2 <- c(50, 13, 5)
lp("max", f.obj2, f.con2, f.ineqs2, f.rhs2)
lp("max", f.obj2, f.con2, f.ineqs2, f.rhs2)$solution
```

##(d)
By working out the final tableau, we can get the Gomory cut.
From (c) we can see the new objective function to be
1.5(x1) + 1(x3) - 0.5(s1) + 2.5(s2) = 7.5
Since all variables are non-negative rounding down the lhs makes the lhs no larger so we get
1(x1) + 1(x3) - 1(s1) + 2(s2) <= 7.5
Now, rounding down the rhs since the lhs is an integer, we get
1(x1) + 1(x3) - 1(s1) + 2(s2) <= 7
This is now the Gomory cut.

##(e)
```{r}
f.obj3 <- c(12, 20, 15, 0, 0, 0, 0, 0)
f.con3 <- matrix(c(2, 5, 3, 1, 0, 0, 0, 0,
                   1, 1, 1, 0, 1, 0, 0, 0,
                   0, 1, 0, 0, 0, -1, 1, 0,
                   1, 0, 1, -1, 2, 0, 0, 1), nrow = 4, byrow = TRUE)
f.ineqs3 <- c("=", "=", '=', '=')
f.rhs3 <- c(50, 13, 5, 7)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)$solution
```

Once I added the Gomory cut, as expected the solution are all integer values, which is what we wanted and it now fits the problem at hand.

##(f)
The optimal solution is x1 = 1, x2 = 6, x3 = 6, and s3 = 1 and the optimal solution is 222.
I was surprised that the optimals had slightly changed, where x1 increased to 1 from 0 and x2 and x3 changed to 6. I imagined that the solution would just be values rounded down to the nearest integer and that the slacks may have increased to compensate for the rounding down of the variables.
The Gomory cut reduced the optimal solution to an integer value which was expected as the optimal without the Gomory cut was 222.5 and for it to be integer we would expect the optimal value to be at 222.



####Theon Coursework####


##(a)
```{r}
RegA <- cbind(t(seq(0, 20, by = 1)), 25)
RegB <- cbind(t(seq(2, 40, by = 2)), 50)
RegC <- seq(3, 60, by = 3)
#Start off by making the sets A, B, and C as described in the coursework

actions <- function(x, A, B, C){
  unionAC <- union(A, C)
  unionABC <- union(unionAC, B)
  #The union function cannot union more than 2 sets at once.
  
  if ((is.element(x, B)) == TRUE){
    A_x <- unionABC[unionABC <= x]
    }
else
  A_x <- unionABC[unionABC < x]
  A__x <- sort(A_x, decreasing = FALSE)
  return (A__x)
}
#This displays the action set values which are less than x but also determines whether to include the value of x if x is a subset of set B.

actions (4, RegA, RegB, RegC)
```

I tried to create a subset of the union set after researching, but I was not able to replicate the code used in the examples online after coding   newdata <- subset(union | x < union).

I then tried to code it in a different way but this did not get far   for (a in union)
    union. <- union[union < x]

I then tried to to make it so my code tells me if it is set B with the following code   if.contains <- x contains(value = x, B)
but it did not work, I think I needed to install a package but I could not find the name.

I then tried     if.contains <- (x %in% B) but an error came up saying I cannot do do as many calculations in one go.

```{r}
V <- function (A, B, C, n, x){
  
  Actions <- actions (x, A, B, C)
  
  for (a in Actions){
     if (a == 0){
      reward <- 0
      #This relates to n not changing by any value
      }
    else {
      reward <- -1
      #This relates to n decreasing by 1 for every dart thrown and the sum of the rewards will denote how may optimal takes it will tak to reach the score.
      }
    
    if (n == 0){
      if (x == 0){
      V_0 <- 0}
      #Here we need to impose the boundary conditions.
      #For n equal to zero and x equal to zero the terminal reward is zero.
      if (x > 0)
      V_0 <- -100
      #For n equal to zero and x greater than zero then we let the terminal reward be negative 100 as this makes it easier when we are finding the maximum number of takes needed to reach the score x. This value is an arbitrarily large number so that it is not considereed with the number of darts taken.
    return ( V_0)}
    else {
      V_n.1 <- V(A, B, C, n - 1, x - a)
      V_n <- reward + V_n.1
      #For n not equal to zero then we find use the optimality equation as shown in the notes.
      list <- list()
      #I want to make an empt list so that the for loop fills up the list
      i <- 1
      list [[i]] <- V_n
      #The possible solutions are then inputted into the list and then unlisted and then the maximum is being found.
      i <- i + 1
      m <- unlist(list)
      Max <- max(m)}
    return (Max)}}


#V(A, B, C, 2, 20)
```
      
      
      }
  }
  return (V_n)}
  
  #for (n in 1:n){
   # for (a in Actions){
   #   V_n.1 <- V(A, B, C, n - 1, x - a)} }
    
   # V <- max(reward[a] + V_n.1 ) }

  

V(A, B, C, 12, 20)


State - A U B U C
Time steps - n time steps
Action - {A__x}
Rewards - fewest number of steps OR minimum number of steps
Evolution - x minus f(A, B, C, n, x)
Value function - 


 for (n in 2:n){
    f. <- f(A, B, C, n - 1)
    n <- n - 1}
  if (n == 1){
    action_final <- actions$A__x[actions$A__x <= B] 
  }}
  #actions$A__x
  
  I tried to code that the last Action set needs to be a subset of B.
  
  
   for (i in Actions){
      }}
    #return (reward)
  
  for (i in 1:n){}
  V_n <- max(R ) 


    for (i in 1:n){
      for (i in Actions){
    if (i == 0){
      reward <- 0}
    else {
      reward < -1
    }}
  }
  V_n <- max(reward[i] ) 
  
  These are my different attempts to try and code the function.
  
  I then kept getting difficulties with reordering the code and whether to put it in which for loop.


I had to comment out the example when using my code as it was taking a very long time to knit and I was not able to figure out why it was taking this long.

I struggled a fair amount during part (b) of this coursework. My though process during this coursework was to find the Action Set which is the set of values which someone can take, given they are at x. After several attempts and reordering in part (b), I coded the optimality equation but it does not seem to be working fully and after shifting the code around as well as adding small pieces of code I kept getting that many of my variables came back as 'object not found'. I eventually managed to output a '-100' which I although know is not correct, it was further than I was originally at. I feel I understand what I am trying to do throughout this process but I cannot put it into a full set of working code.
