---
title: "Anaxagoras_and_Zeno"
author: "Kieran Patel"
date: "02/11/2017"
output: html_document
---

```{r}
library(lpSolve)
```




###Anaxagoras Coursework###

##(i)
```{r}
f.obj1 <- c(1, 1)
f.con1 <- matrix(c(1, 2, 3, 1), nrow = 2, byrow = TRUE)
f.ineqs1 <- c("<=", "<=")
f.rhs1 <- c(100, 75)
lp("max", f.obj1, f.con1, f.ineqs1, f.rhs1)
lp("max", f.obj1, f.con1, f.ineqs1, f.rhs1)$solution
```
The optimal solution is 10 chairs and 45 tables and the optimal solution is 55.



##(ii)
We augment the problem, as seen in the lecture notes.
```{r}
f.obj2 <- c(1, 1, 0, 0)
#Where the 3th and 4th column are the coefficients of the slack variables s1 and s2
f.con2 <- matrix(c(1, 2, 1, 0, 3, 1, 0, 1), nrow = 2, byrow = TRUE)
#Where the 3rd, 4th, 5th and 6th column are the coefficients of the slack variables s1 and s2
f.ineqs2 <- c("=", "=")
f.rhs2 <- c(100, 75)
lp("max", f.obj2, f.con2, f.ineqs2, f.rhs2)
lp("max", f.obj2, f.con2, f.ineqs2, f.rhs2)$solution
```

I initially, wrote 'obj.2 <- c(1, 1)' which returned 'Error: no feasible solution found[1] 0 0'. I realised that I should have typed 'obj.2 <- c(1, 1, 0, 0)' as the slack variables have coefficients of zero in the objective function.
This then returned the following response 'Success: the objective function is 1e+30 
[1] 1e+30 0e+00 0e+00 0e+00' which I then realised was wrong as I did not include all of the constraints and had written the constraints wrong so I changed 'con.2 <- matrix(c(1, 2, 3, 1, 1, 1), nrow = 2, byrow = TRUE)' to 'con.2 <- matrix(c(1, 2, 1, 0, 3, 1, 0, 1), nrow = 2, byrow = TRUE)'.

The optimum is the same as in (i) so it is verified. The slack variables both take zero at the optimum. We could have worked this out previously since if the optimal values of the chairs at 10 and tables at 45 are substituted into the constraints we can see the inequality hold and is tight so we already know the slack variables are zero.



##(iii)
f.obj3 <- c(1)
f.con3 <- matrix(c(1, 3, 1), nrow = 1, byrow = TRUE)
f.ineqs3 <- c("=", "=")
f.rhs3 <- c(150)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)$solution


obj.3 <- 1
con.3 <- c(1, 3, 1)
ineqs.3 <- "="
rhs.3 <- 150
lp("max", obj.3, con.3, ineqs.3, rhs.3)
lp("max", obj.3, con.3, ineqs.3, rhs.3)$solution


I did not realise I was supposed to add the constraint to the problem already in place in (ii).
```{r}
f.obj3 <- c(1, 1, 0, 0, 0)
f.con3 <- matrix(c(1, 2, 1, 0, 0, 3, 1, 0, 1, 0, 1, 3, 0, 0, 1), nrow = 3, byrow = TRUE)
f.ineqs3 <- c("=", "=", "=")
f.rhs3 <- c(100, 75, 150)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)
lp("max", f.obj3, f.con3, f.ineqs3, f.rhs3)$solution
```

I simply, added a new constraint to the problem in (ii), it appears the constraint already in place from (ii) are stronger than the new constraint which is why the optimal solution for profit remains at the same value.



##(iv)
```{r}
f.obj4 <- c(1, 1, 0, 0)
x <- 99
f.con4 <- matrix(c(1, 2, 1, 0, 3, 1, 0, 1), nrow = 2, byrow = TRUE)
f.ineqs4 <- c("=", "=")
  repeat{
    f.rhs4 <- c(x, 75)
    print(lp("max", f.obj4, f.con4, f.ineqs4, f.rhs4))
    print(lp("max", f.obj4, f.con4, f.ineqs4, f.rhs4)$solution)
    x <- x - 1
    if (x == 0){
      break}
  }
```

I initially coded
lp("max", f.obj41, f.con41, f.ineqs41, f.rhs41)
lp("max", f.obj41, f.con41, f.ineqs41, f.rhs41)$ solution but the output was not printing out so I changed it to after researching how to use repeat loops in the internet.
print(lp("max", f.obj41, f.con41, f.ineqs41, f.rhs41))
print(lp("max", f.obj41, f.con41, f.ineqs41, f.rhs41)$solution)


The optimum value decreases as the value of b1 decreases. The optimal basis changes for each value of b1 too where x1 increases by 0.2 up to x1 = 25 which is a tight condition and then it decreases by 1 for each decrease of 1 to b1. The value of x2 decreases by 0.6 for each single decrease in b1. The slack variable s1 does not change as b1 changes from 99 to 1, but s2 does begin to increase by 1 once the optimum value passes 25.



##(v)

f.obj5 <- c(1, 1, 0, 0)
x <- -1
f.con5 <- matrix(c(1, 2, 1, 0, 3, 1, 0, 1), nrow = 2, byrow = TRUE)
f.ineqs5 <- c("=", "=")
  repeat{
    x <- x + 1
    f.rhs5 <- c(x, 75)
    ; if (.Last.value == lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)) break; print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5))
    print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution)}


I tried this as my code and had to terminate the output as it kept increasing to infinity and was at x2 = 75 and s1 = 1216 so this was not looking useful or correct.

I am trying to create a loop which keeps increasing b1 by 1 until the optimal basis from the b1 = i - 1 is the same as the b1 = i and then I would like the output for this final solution to be printed.


f.obj5 <- c(1, 1)
x <- -1
f.con5 <- matrix(c(1, 2, 3, 1), nrow = 2, byrow = TRUE)
f.ineqs5 <- c("=", "=")
  repeat{
    x <- x + 1
    f.rhs5 <- c(x, 75)
    ; if (.Last.value == lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)) break; print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5))
    print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution)}


I then try and repeat the code but took out the slack variables as the output for the previous code I wrote kept increasing the slacks.



f.obj5 <- c(1, 1, 0, 0)
x <- -1
f.con5 <- matrix(c(1, 2, 1, 0, 3, 1, 0, 1), nrow = 2, byrow = TRUE)
f.ineqs5 <- c("=", "=")
  repeat{
    x <- x + 1
    f.rhs5 <- c(x, 75)
    ; if (a == lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution) break; print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5))
    a <- lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution
    print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution)}


I then realised that I was checking the condition against the optimum value as opposed to the optimal basis, but this still does not return a relevant answer.


f.obj5 <- c(1, 1)
x <- -1
f.con5 <- matrix(c(1, 2, 3, 1), nrow = 2, byrow = TRUE)
f.ineqs5 <- c("=", "=")
  repeat{
    x <- x + 1
    f.rhs5 <- c(x, 75)
    ; if (.Last.value == lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution) break; print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5))
    print(lp("max", f.obj5, f.con5, f.ineqs5, f.rhs5)$solution)}


I then thought of removing the slacks to provide a more loose constraint but the same issue reappears. I was not able to progress further than this and given (vi) requires the data found from (v), I was not able to progress.




###Zeno Coursework###


```{r}
bfs <- c(6, 4, 3)
c <- c(2, 3)
con <- rbind(c(1, 2), c(1, 1), c(1, 0))
con
```

I have used application 2.3 as an example for this coursework Zeno.



##(ai)
I needed to code the number of constraints and then add these many slacks with coefficients of zero and after researching the repeat function, I found I can repeat zero a variable number of times. The nrow function sum the number of rows the matrix of constraints holds.

I initially thought I was meant to augment the problem as we did during chapter 1 when we add the slack variables and the following code shows my attempt.

y <- nrow(con)
S <- rep(0, y)
#S represents the slacks
A <- cbind(con, S)
A


I attempted to augment the problem so it looks like it does in the initial tableau.
y <- nrow(con)
y
S1 <- rep(0, y)
S2 <- rep(0, y)
S3 <- rep(0, y)
A <- cbind(con, S1, S2, S3)
A


But the rows did not show an identity matrix in the slack rows. So I then researched the functions given in the hints section for Zeno and diag was the relevant function here.


```{r}
x <- length(c)
y <- nrow(con)
I <- diag(x = 1, nrow = y, ncol = y)
s <- rep(0, y)
c
C <- c(c, s)
C
```



##(aii)
```{r}
A <- cbind(con, I)
A. <- cbind(con, I, bfs)
A.
S <- rep(0, y)
#Finding where the initial basis lies which is between x+1 and x+y
X <- x + 1
Y <- x + y
B <- A[,c(X : Y)]
B
I.B <- c(X : Y)
I.B

cB <- C[c(X : Y)]
cB
bfs. <- x + y + 1
BFS <- A.[,c(bfs.)]
BFS
```

I tried this code initially but I kept getting errors
I.B <- A[,c(x + 1 : x + y)]
So I changed the input to separate variables, thinking the code had too much input to calculate at once. I used this example to determine what to do
z <- diag (3)
z[, c(2:3)]

I then needed to find the coefficients of the initial basis so I needed to add on the coefficients of the slacks onto the c, the objective function. I found the code C <- cbind(c, s) did not work as the cbind function work for matrices only so I changed cbind to c.



##(b)

I am trying to rename the columns in the matrix so that they can be extracted in a way that once it is, it is more clear than column numbers. This was not working well I decided to remain with column names whereby the initial basis is the slacks which will begin from the Xth column (since X = x + 1)

I tried renaming column names so that they read x1, x2,..., s1, s2,... using the following 
q <- c('x', 1:x)
w <- c('s', X:Y)
r
colnames(A) <- paste(r)
colnames(A) <- paste('x',1:x)

This came up with errors so I then thought I would try and research into strings but this did not get anywhere.

```{r}
A.
Name <- c(1: Y)
Name
colnames (A.)[1:Y] <- paste (Name)
A.
#We now add cB and Basis to our matrix
C. <- c("", "", C, "")
#While it is not necessary to add this row onto the matrix, it helps with presentation
C.
A.. <- cbind(cB, I.B, A.)
A... <- rbind(C., A..)
A..
A...
```

I decided to have another attempt to rename the columns for the variables. I initially used
Name <- c(1: Y)
Name
names(A) <- Name
colnames (A) <- paste (Name, 1:Y)

and had to fiddle around which of these naming functions worked as I was getting an output that renames columns with '1 1' as opposed to a '1'. I eventually managed to fix it but tweaking the code and using one renaming method while turning the other into text for the time being with the #.


I initially tried cB_T.M <- cB*A
cB_T.M
colSums (cB_T.M, na.rm = FALSE, dims = 1)


While this works the same, the new code which I took from the hint sheet in the lecture notes mentions the use of %*% and it makes the code more compact
```{r}
cB_T.M <- (t(cB)) %*% (A)
cB_T.M
cB_T.BFS <- (t(cB))%*%BFS
cB_T.BFS
cB_T.M. <- cbind("Objective Row", "", cB_T.M, cB_T.BFS)
cB_T.M.
c__cB_T.M <- C - cB_T.M[1:Y]
c__cB_T.M
c__cB_T.M. <- cbind("Simplex Coefficients", "", t(c__cB_T.M), "")
c__cB_T.M.
A.... <- rbind (A..., cB_T.M., c__cB_T.M.)
A.
A..
A...
A....

max <- which.max(t(c__cB_T.M))
max
```

For some reason c__cB_T.M. <- cbind("Simplex Coefficients", "", c__cB_T.M, "")
kept giving back c__cB_T.M. as a matrix as opposed to a row vector like cB_T.M., I then decided to transpose the c__cB_T.M vector and it managed to work.


##(c)
```{r}
#We know fro b what the given column is which is named max. The given column  can only be between the 3rd and 2+Y th columns but I am using an extrapolated version of the tableau, which is M in the notes so that if any columns which are not the variables are selected, an error message is displayed about the dimensions not matching eg. if var <- .A.[,6] we are shown 'Error in .A.[, 6] : subscript out of bounds'.

.A. <- A.[, 1 : Y]
.A.
var <- A.[,max]
Ratios <- BFS / var
Ratios
minpositive = function(x) min(x[x > 0])
minpositive(Ratios)
match (minpositive(Ratios), Ratios)
```

var <- .A.[,3]
This was my initial code for the given column but I realised that the given column depends on the maximum value of the objective 

I was then finding the ratios to find the largest of these so we can compute which is the pivot row.

The issue with using which.min is that this function has limitations in that the minimum may be negative or zero but we need the smallest nonzero value. The current code has fewer limitations, however, 'inf' may be shown if all ratios are all infinity.

The match function returns the row for which the minimum positive value of the ratios relate to and this row is the pivot row. I knew of the match function after similarly working with Excel and understanding its match function which returns the cell coordinates for the first match of the conditions.